\begin{abstract}

    My thesis explores the domain of automatic commit message generation with deep learning methods. The results achieved in deep learning have increased significantly in the last six years. This progress is especially noticeable in natural language processing (\textbf{NLP}). The generation of commit messages can be viewed as a neural machine translation task in NLP\@. It involves translating code into natural language to describe the changes made in the commit. I claim that considering the current progress in NLP, it is possible to generate descriptive commit messages from code modifications. 

    My work consists of a literature review, data collection and analysis, deep learning models training, evaluation, and discussion of results. Regarding model training, I experimented with models of different sizes, an additional module to solve the problem with message style adaption, and an additional module to solve the problem of limited context. During the literature review, metrics analysis is conducted to identify the most suitable metrics to generate commit messages. I also carried out the experiment to prove that my metrics make sense and correctly identify how good the generated message is. 
    
    In summary, my thesis includes an analysis of the problem from a data, metrics, and model perspective. I also carried out experiments to reproduce the results from other papers. In addition, I proposed an idea to mitigate the problem of handling large commits. At the end of my work, I evaluated all trained models and came to a conclusion about the current state of commit messages generation.
\end{abstract}